Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-08-08 Ben_Manuscripts/hdphmm/Draft.tex

  In this work, we are particularly interested in the design of lyotropic liquid 
  crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
  phases can be cross-linked into mechanically strong membranes capable of highly 
  selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
  are characterized by hexagonally packed, uniform-sized and straight pores, an ideal
  geometry for high throughput transport. Their pores are lined with the LLC monomer's
  functional groups which can potentially be designed to interact with solutes in a 
  chemically-specific manner. It would be extremely useful to experimentalists if we 
  could connect LLC monomer design with measurable quantities. This goal is complicated
  by the non-Brownian hopping and trapping behavior induced by the membrane's diverse
  topology and inhomogeneous structure resulting in subdiffusive behavior out to 
  multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}

  We have already advanced our understanding of LLC monomer design in our previous work
  by building stochastic time series models based on the chemical intuition gained
  from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
  membrane.~\cite{coscia_chemically_2019,coscia_capturing_2020} We identified three 
  classes of trapping behavior which result in subdiffusion out to the microsecond 
  time scale. In our first approach to stochastically model this behavior, we treated 
  the system's dynamics as a sequence of anti-correlated hops between power law distributed periods of 
  entrapment, a framework called subordinated fractional Brownian or L\'evy
  motion.~\cite{thiel_weak_2014,teuerle_modeling_2013} In a second approach, we treated 
  solute motion as a Markov state model~\cite{pande_everything_2010} with state-dependent
  dynamics, where we parameterized the state transition probabilities between each of
  eight discrete states, defined by the observed trapping mechanisms, as well as the 
  solute dynamics within each of these states. We showed how one could use realizations
  of any stochastic model like these in order to predict macroscopic flux and selectivity.

  Although powerful and insightful, our previous work required considerable human effort
  in order to analyze the MD trajectories and to hypothesize stochastic models which
  matched the behavior of solutes. It would be significantly more useful and broadly 
  applicable if we could design an approach which distinguishes and parameterizes varied
  solute dynamics in a data-driven way and leaves a facile way to generate an ensemble
  of stochastic trajectory realizations that can be used to predict macroscopic transport
  properties.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10\% water by weight, totaling 62,000 atoms. Approximately one third of 
  the water molecules occupy the tail region with the rest are near the pore 
  center.~\cite{coscia_chemically_2019}

>>>>

  In this work, we are particularly interested in the design of lyotropic liquid 
  crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
  phases can be cross-linked into mechanically strong membranes capable of highly 
  selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
  are characterized by hexagonally packed, uniform-sized and straight pores, an ideal
  geometry for high throughput transport. Their pores are lined with the LLC monomer's
  functional groups which can potentially be designed to interact with solutes in a 
  chemically-specific manner. It would be extremely useful to experimentalists if we 
  could connect LLC monomer design with measurable quantities such as diffusivity and
  selectivity. This goal is complicated by the non-Brownian hopping and trapping behavior
  induced by the membrane's diverse topology and inhomogeneous structure resulting in 
  subdiffusive behavior out to multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}

  %r1_comment2
  A number of approaches have been explored to inexpensively extend the relatively short
  trajectories output by molecular simulations which exhibit complex diffusive behaviors
  onto much longer timescales. The trajectory extending kinetic Monte Carlo (TEKMC) approach
  divides a unit cell into a three dimensional grid and quantifies the transition 
  probabilities of particles that can move between each subdivision allowing one to generate
  stochastic trajectories that reach the diffusive regime~\cite{neyertz_trajectory-extending_2010}
  It is possible to parameterize TEKMC models with relatively short simulation trajectories 
  which facilitates systematic studies of small perturbations to material constituents.~\cite{hanson_computer_2012}
  In a related approach, dynamic bond percolation (DBP) theory can be used to quantify the
  rate of transition between sites in a dynamically disordered medium and long timescale
  behavior can be propagated using kinetic Monte Carlo simulations.~\cite{druger_dynamic_1983}
  Webb et al. extended this approach by defining sites based on the molecular details of 
  a polymer electrolyte system, thus allowing them to explain ion diffusivities predicted by
  the model in terms of chemically specific interactions.~\cite{webb_chemically_2015}

%  We have already advanced our understanding of LLC monomer design in our previous work
%  by building stochastic time series models based on the chemical intuition gained
%  from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
%  membrane.~\cite{coscia_chemically_2019,coscia_capturing_2020} 
  Based on the chemical intuition gained from our previous qualitative MD studies of 
  solute transport in an H\textsubscript{II} phase LLC membrane, we have contributed our
  own stochastic modeling approaches with goals similar to both TEKMC and DBP.~\cite{coscia_chemically_2019,coscia_capturing_2020}
  We identified three classes of trapping behavior which result in subdiffusion out to the microsecond 
  time scale. In our first approach to stochastically model this behavior, we treated 
  the system's dynamics as a sequence of anti-correlated hops between power law distributed periods of 
  entrapment, a framework called subordinated fractional Brownian or L\'evy
  motion.~\cite{thiel_weak_2014,teuerle_modeling_2013} In a second approach, we treated 
  solute motion as a Markov state model~\cite{pande_everything_2010} with state-dependent
  dynamics, where we parameterized the state transition probabilities between each of
  eight discrete states, defined by the observed trapping mechanisms, as well as the 
  solute dynamics within each of these states. We showed how one could use realizations
  of any stochastic model like these in order to predict macroscopic flux and selectivity.

  Although powerful and insightful, TEKMC, DBP and our previous work required considerable
  human effort in order to analyze the MD trajectories and to hypothesize stochastic models 
  which matched the behavior of solutes. It would be significantly more useful and broadly 
  applicable if we could design an approach which distinguishes and parameterizes varied
  solute dynamics in a data-driven way and leaves a facile way to generate an ensemble
  of stochastic trajectory realizations that can be used to predict macroscopic transport
  properties.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using our
  previously developed atomistic molecular model of four pores in a $9.4\times9.4\times7.6$ nm 
  monoclinic unit cell with 10\% water by weight, totaling 62,000 atoms. Approximately
  one third of the water molecules occupy the tail region with the rest are near the
  pore center.~\cite{coscia_chemically_2019}

  %r1_comment3  
  We modeled water molecules using the TIP3P water model,~\cite{jorgensen_comparison_1983}
  We used the Antechamber module shipped with AmberTools18 to parameterize monomers and 
  solutes using the Generalized Amber Force Field (GAFF).~\cite{wang_development_2004} We 
  assigned charges using the AM1-BCC model.~\cite{jakalian_fast_2000,jakalian_fast_2002}
  GAFF is very commonly used in simulations with diverse organic chemical functionality, 
  and was developed for consistency within the AMBER ecosystem, which uses TIP3P as the
  water model. We chose these standard force fields and generalizable parameterization 
  approaches so that a range of solutes and monomers could be self-consistently and easily 
  studied. The main goal of this work is to develop stochastic models which can reproduce
  solute behavior on MD time scales, independent of the choice of force field. In future 
  work, it may be beneficial to employ specialized force fields, including polarizable 
  force fields or other higher-order effects in order to improve the accuracy of structure 
  parameters and selectivity predictions.
